Reproducible Research: Peer Assessment 1
========================================

## Loading and preprocessing the data
```{r}
library(ggplot2)
library(plyr)
data <- read.csv("activity.csv")
dailytotals <- aggregate(data$steps, by=list(data$date), FUN=sum)
colnames(dailytotals) <- c("Date","Steps")
dailytotals$Date <- as.Date(dailytotals$Date)
```

## What is mean total number of steps taken per day?

First, we'll look at a histogram of the total number of steps per day through the two-month period covered by the data set.

```{r fig.height=4}
par(las=2)
ggplot(dailytotals, aes( x = factor(Date), y = Steps)) + geom_histogram(stat= "identity") + theme(axis.text.x = element_text(size=8, angle=45, hjust=1))
```

```{r echo=FALSE}
good <- complete.cases(dailytotals)
totals <- dailytotals[good,]
meanDaily <- mean(totals$Steps)
medianDaily <- median(totals$Steps)
```
For this two-month period, the mean number of steps is `r meanDaily` and the median number of steps is `r medianDaily`.


## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
